---
title: "Concreteness Baseline"
author: "Mark Angelo Gabriel  - 22210681"
date: "15/03/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r}
library(MASS)
library(ggplot2)
library(e1071)
```

# Data setup

```{r}
df <- read.csv('../data/Concrete_DayZero_15032021.csv')
```

With this analysis, we want to be able to create a model that predicts concreteness with data that can be obtained at day zero.
Remove unnecessary variables for preliminary analysis. Age has also been removed in favor of Age.Group. 

```{r}
df <- within(df, rm("Response.x", "Steps", "Effective", "Emotion", "invalid_response", "Age.r"))
df$IsolationDays.0 <- as.numeric(df$IsolationDays.0)
df <- na.omit(df)
summary(df)
```

We want to see how well the available variables can predict `Concrete`. For this, we'll use linear regression.

But first, we'll be setting up the train/test split.

```{r}
set.seed(5014)
test.idx <- sample(nrow(df), floor(nrow(df)*0.1))
df.train <- df[-test.idx,]
df.test <- df[test.idx,]

c(nrow(df.train), nrow(df.test))
```

# Linear Regression Models

Since we have a lot of variables and we estimate only a handful being significant, we'll use stepwise selection to create our predictor model. 

```{r}
concrete.lm <- lm(Concrete ~ ., df.train)
concrete.ss <- lm(Concrete ~ 1, df.train)
concrete.ss <- step(concrete.ss, scope=list(upper=concrete.lm), trace=0)
summary(concrete.ss)
```
The adjusted R-squared score on the `Concrete` model is 0.04031. This is an extremely low score which suggests that our current model is inadequate for accounting for the variance of `Concrete`.

# Testing accuracy

```{r}
concrete.pred.train <- predict(concrete.ss)
concrete.pred.test <- predict(concrete.ss, newdata=df.test, type="response")

# RMSE calculation
sqrt(mean((concrete.pred.train-df.train$Concrete)^2))
sqrt(mean((concrete.pred.test-df.test$Concrete)^2))
```

```{r}
boxplot(df$Concrete)
```

```{r}
df.test$ConcretePred <- concrete.pred.test
df.test <- df.test[order(df.test$Concrete),]

x.label <- 1:dim(df.test)[1]

ggplot(df.test, aes(x=x.label)) + 
  geom_line(aes(y=Concrete, color="True")) + 
  geom_line(aes(y=ConcretePred, color="Predicted")) +
  theme_classic()
```

